Demand Forecasting of individual Probability Density Functions with Machine Learning

### 2020-08-19 LaTex/paper.tex
reviewed Cyclic Boosting section

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e. the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g. optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e. lost sales) and waste rate. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.

So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the calculation of the optimal order quantity. Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model. This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case. And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level. Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method. Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand. Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth. It should be noted that these parameters are not independent between products. For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically. Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance. The features may or may not differ between the mean and variance estimation models. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD models using Cyclic Boosting.

@@

Cyclic Boosting \cite{Wick2019} is a type of generalized additive models using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters. Major benefits of Cyclic Boosting are its accuracy, performance even at large scale and providing fully explainable predictions which are of vital importance in practical applications, in particular for multi-national retail vendors.

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all target values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example the sales of a specific SKU along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges. During training,  regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a  specific feature contributes to each individual prediction.

@@

When looking at the demand of individual SKUs, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.

>>>>

Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision. More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g. \cite{silver1998} for an overview. In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$. In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g. \cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}. Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g. a business day). For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly. However, it should be noted that demand is not independent between different samples for a given date. For example, a promotion applied to one product can affect the sales of related products within the assortment. This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.

The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g. in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}. It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e. the order quantity) directly. An overview of a range of different approaches can also be found in \cite{huber2019data}. Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages. First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g. optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e. lost sales) and waste rate. From a practical perspective, separating the demand forecast from the operational decisions (i.e. calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand. It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer. From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other. Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e. the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}. Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.

So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution. The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g. \cite{alwan2016}) or via sample average approximation (SAA), see e.g. \cite{shapiro2014} for an overview. However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price, advertisements, or weather influence, and therefore are highly correlated with, the demand. Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e. the calculation of the optimal order quantity. Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately. Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model. This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case. And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level. Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method. Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.

@@

To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted. As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand. Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance. At least in principle, any method can be used. However, as discussed in sec. \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches. Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm. This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance, and the features may or may not differ between the mean and variance estimation models. The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.

In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD using two Cyclic Boosting models.

@@

Cyclic Boosting is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters. Major benefits of Cyclic Boosting are its accuracy, performance, even at large scale, and providing fully explainable predictions, which are of vital importance in practical applications.

The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature. The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data. Single data records, for example the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
The individual predictions $\hat{y_i}$  can then be calculated as:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges. During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm. The deviation of each factor from $f^k_j=1$ can then be used to explain how a  specific feature contributes to each individual prediction.

@@

When looking at the demand of individual product-location-date combinations, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.
